Peer-to-Peer Permissionless Consensus via Reputation

### 2023-05-12 freechains/freechains.tex
...

\begin{table}
\centering
\includegraphics[width=0.75\textwidth]{general.png}
\caption{General reputation operations in public forums.}
\label{fig.general}
\end{table}

We propose a reputation system in which users can spend tokens named \reps to
post and rate content in the forums:
a \code{post} initially penalizes authors until it consolidates and counts
positively;
a \code{like} is a positive feedback that helps subscribers to distinguish
content amid excess;
a \code{dislike} is a negative feedback that revokes content when crossing a
threshold.
Table~\ref{fig.general} summarizes the reputation operations and their goals.
To prevent Sybils, users with no \reps cannot perform these operations,
requiring a welcoming like from any other user already in the system.
The fact that likes are zero-sum operations, which only transfer reputation,
eliminates the incentives from malicious users to invite Sybils into the
system.
The only way to generate new \reps is to post content that other users
tolerate, which demands non-trivial work resistant to automation.%
\footnote{
    Note that to some extent AI content may be indistinguishable from humans,
    but its frequency and relevance in forums are still subject to human
    evaluation in our proposal.
}

@@

\subsection{Basic Reputation Consensus}

@@

Our key idea to stablish consensus in forum DAGs is to favor forks with posts
from users that constitute the majority of the reputation in the network.
These forks have more associated work from active users and are analogous to
longest chains in Bitcoin.
%
In technical terms, we can adapt a topological sorting algorithm to favor
reputation when deciding between branches in a forum DAG.

Figure~\ref{fig.reps}.A illustrates the evolution of a forum DAG in the
presence of a network partition:
A common prefix has signed posts from users $a$, $b$, and $c$, when they were
still all connected.
Then, user $c$ disconnects from $a$ and $b$, and evolves in separate along with
new users $x$ and $y$ with posts in \code{branch-1}.
In the meantime, $a$ and $b$ participate with posts in \code{branch-2}, but
with less activity.
After some time, the partitions reconnect and new posts from $x$, $b$ and $c$
merge the branches with a common suffix.

Even in the ordinary case of a legit temporary network partition, note that the
two branches may still show conflicts when merging.
For instance, posts with dislike operations in one branch could possibly block
posts from users in the other branch.
Therefore, we need to order the branches with a global consensus, such that we
can validate operations consistently across the network.
%
Note also that, to reach consensus, we should preferably rely on the maximum
information that all branches agree on, which is exactly the posts in the
common prefix.

In Figure~\ref{fig.reps}.A, we assume that within the common prefix, users $a$
and $b$ have contributed with better content and therefore have more reputation
combined than $c$ has alone (i.e., $8+5>3$).
%
Based on that information, Figure~\ref{fig.reps}.B indicates the resulting
consensus order between posts in the forum DAG:
    the common prefix comes first,
    followed by all posts in \code{branch-2},
    followed by all posts in \code{branch-1},
    followed by the common suffix.
%
The \code{branch-2} with users $a$ and $b$ take priority over \code{branch-1}
with user $c$ because, before the forking point, $a$ and $b$ have more
reputation than $c$, $x$, and $y$ combined.

At any point in the consensus timeline, if an operation fails, all remaining
posts in the offending branch are removed from the primary DAG.
As an example, suppose that the last post by $a$ (in gray) is a dislike to user
$c$.
Then, it's possible that the last post by $c$ (in red), now (suppose) with
\nreps{0}, is rejected together with all posts in sequence, including those in
the common suffix.
%
This rejection rule for remaining posts is a direct consequence of tamper proof
Merkle~DAGs~\cite{p2p.ipfs,p2p.bitcoin} typically used in permissionless
protocols, which cannot support graph modifications.

The basic consensus rule is therefore straightforward:
Whenever a fork is found, we create sets of users with posts in each branch
($B1=\{c,x,y\}$ and $B2=\{a,b\}$).
Then, we take each set, and sum its users \reps based on the common prefix
($S1=3$ and $S2=13$), since it is the maximum information that all branches
agree.
Finally, we order the branches from the highest sum of \reps found
(Figure~\ref{fig.reps}.B).
While applying the branches in order, if any post operation fails, all
remaining posts are rejected, as if they never existed.

There are some other relevant considerations about forks and merges:
%
Peers that first received branches with less reputation will need to reorder
all posts starting at the forking point.
This might involve removing content in the end-user software.
This behavior is similar to Bitcoin's blockchain reorganization, when a peer
detects a new longest chain. % and disconsiders old blocks.
%
Likewise, peers that first saw branches with more reputation just need to put
the other branch in sequence with no reordering at all.
%and do not need to recompute anything.
This behavior is expected to happen in the majority of the connected network.
%
Note that unlike Bitcoin, forks are not only permitted but encouraged due to
the local-first software principle~\cite{p2p.local}, in which networked
applications can work locally while offline.
However, the longer a peer remains disconnected, the more conflicting
operations it may see, and the higher are the chances of reordering when
rejoining.

In summary, it is important to remark that
    (i) branches always introduce conflicts, which
    (ii) require consensus to order operations, which
    (iii) is based on reputation of an agreed common prefix,
    (iv) leading to a global and deterministic and result.
%
Note that the consensus order exists only for accountability purposes, and is
a view of the DAG structure, which remains as the primary source of data.

\subsection{Malicious Behaviors and Hard Forks}

Now, consider a malicious scenario in which user $c$ in Figure~\ref{fig.reps}
intentionally disconnected from the network for weeks to cultivate fake
identities $x$ and $y$ and accumulate \reps.
When merging, these fake users become legit and now constitute the majority of
the forum in terms of reputation.
%
However, because the common prefix is the basis of consensus, users in the
branch with more reputation can still react even after the attack.
For instance, users $a$ and $b$ can pretend that they did not yet see
\code{branch-1} and post extra dislikes to user $c$ from \code{branch-2} before
the common suffix so that a further merge removes all posts of \code{branch-1}
from the DAG.
%
The example shows that this kind of attack is innocuous because it requires at
least half of the \reps accumulated in the common prefix.
%
The same countermeasure can be used against a minority of users colluding to
boost or protect their own posts.

@@

As a counterpoint, maybe users $a$ and $b$ have abandoned the forum for months,
and thus \code{branch-1} is actually legit.
In this case, users $a$ and $b$ might be the ones trying to take over the
forum.
Yet another possibility is that both branches are legit but became disconnected
for a long period of time.
In any case, it is unacceptable for users that a very old remote branch affects
a long active forum.
%
For this reason, the consensus algorithm includes an extra constraint that
prevents long-lasting local branches to merge, which leads to \emph{hard forks}
in the network.
A hard fork occurs when a local branch crosses a predetermined and irreversible
threshold of \emph{7 days} or \emph{100 posts} of activity.
In this case, regardless of the remote branch reputation, the local branch
takes priority and is ordered first.
This situation is analogous to a hard fork in Bitcoin and the branches will
never synchronize again.
More than simple numeric disputes, a hard fork represents a social conflict in
which reconciling branches is no longer possible.
%
Figure~\ref{fig.hard} illustrates hard forks by distinguishing \emph{stable
consensus}, which cannot be reordered, from \emph{unstable consensus}, which
may still be affected by incoming branches.
The activity threshold counts backwards, starting from the newest local post
in the unstable consensus to older posts until they are permanently frozen.

In summary, the rules to merge a branch from a remote machine \code{j} into a
branch from a local machine \code{i} are as follows:
\begin{itemize}
    \item \code{i} is first if it crosses the activity threshold of
          \emph{7 days} or \emph{100 posts}, regardless of \code{j}.
    \item \code{i} or \code{j} is first, whichever has more reputation in the
          common prefix.
    \item otherwise, branches are ordered by an arbitrary criteria, such as
          lexicographical order of the post hashes immediately after the
          common prefix.
\end{itemize}

To conclude this section, we should keep in mind that our consensus algorithm
is based on subjective evaluations of users and posts.
For this reason, unlike Bitcoin, in which forks are discouraged and a single
blockchain must prevail, public forums can split and diverge indiscriminately.
For instance, users can explicitly apply hard forks at any point in time or
``reboot'' a community from a previous state.
Ultimately, the consensus algorithm provides a transparent mechanism to help
users understand the evolution of public forums and act accordingly.

@@

Finally, we consider that revoking posts is fundamental in the context of
social content publishing, and thus, an important contribution of this work.
As described in Figure~\ref{fig.state}, a forum post has three possible states:
\code{BLOCKED}, \code{ACCEPTED}, or \code{REVOKED}.
%
If the author has reputation, a new post is immediately \code{ACCEPTED} in the
forum.
%
Otherwise, it is \code{BLOCKED} and requires a like from another user.
Blocked posts are not considered part of the forum DAG in the sense that new
posts do not link back to it.
In addition, peers are not required to hold blocked posts and neither
retransmit them to other peers.
However, if blocked posts are not disseminated, new users will never have the
chance to be welcomed with a like.
A reasonable policy is to hold blocked posts in a temporary bag and retransmit
them for some visibility in the network.
%
Once accepted, a post becomes part of the forum and can never be removed
again, since Merkle~DAGs are immutable by design.
%Note that blocked posts that become accepted are always succeeded by a
%\emph{like} (Figure~\ref{fig.forum}).
%
However, if the number of dislikes exceeds an arbitrary threshold (e.g., more
dislike than likes), the post becomes \code{REVOKED} and its payload is not
retransmitted to other peers.
Note that a post hash does not depend on its associated payload, but only on
the payload hash.
Hence, it is safe to remove the payload as long as one can prove its revoked
state.
Later, if the post receives new likes, it means that the payload is still known
somewhere and peers can request it when synchronizing again.

\subsection{Synchronization and Byzantine Peers}

Similarly to Bitcoin, the underlying network protocol replicates on each peer
the full state of the Merkle-DAG representing the causal relationships between
the messages.
When peers synchronize, they mutually synchronize missing branches such that
they reach the same DAG structure~\cite{p2p.byz}.
As they synchronize, each peer runs the consensus algorithm locally and
verifies if all received posts are consistent.

Because the protocol relies on a tamper-proof DAGs, and because each peer
self-verifies the posts as they synchronize, the protocol becomes tolerant to
Byzantine-faulty peers~\cite{lamport.byz}.
%
No information coming from any peer is trusted a priori.
%
For instance, a malicious peer that tries to synchronize branches with erratic
data can be detected on the first inconsistent post.
This allows the correct peer to close the connection immediately and blacklist
the malicious peer from future connections.

Therefore, although Byzantine peers may still perform some kind of denial of
service attacks in the network, they cannot modify the DAG structure, create
new content (without performing work), or impede that two correct nodes
synchronize directly.

@@

We integrated the proposed reputation and consensus mechanism with \FC to
support content moderation and mitigate abuse.
Table~\ref{fig.rules} details the rules we conceived, which are discussed as
follows through a concrete example.
Authors have to sign posts in order to be accounted by the reputation system
and operate in the chains.
We start by creating an identity whose public key is assigned as the pioneer in
a chain named \emph{forum}:

{\scriptsize
\begin{verbatim}
 > freechains keys pubpvt 'pioneer-password'
 4B56AD.. DA3B5F..  <-- public and private keys
 > freechains 'forum' join '4B56AD..'
 10AE3E..           <-- hash representing the chain
 > freechains 'forum' post --sign='DA3B5F..' \
    'The purpose of this chain is...'
 1_CC2184..         <-- hash representing the post
\end{verbatim}
}

The \code{join} command in rule~\code{1.a} bootstraps a public chain,
assigning \nreps{30} equally distributed between an arbitrary number of
pioneers indicated through their public keys.
The pioneers shape the initial culture of the chain with the first posts and
likes, while they gradually transfer \reps to other authors, which also
transfer to other authors, expanding the community.
%
The \code{post} command in sequence is signed by the single pioneer (in this
example) and indicates the purpose of the chain for future users.

The most basic concern in public forums is to resist Sybils abusing the
chains.
Fully peer-to-peer systems cannot rely on logins or {\scriptsize CAPTCHAs}
due to the lack of a central authority.
Viable alternatives include (i) building social trust graphs, in which users
already in the community vouch for new users, or (ii) imposing explicit costs
for new posts, such as proof of work.
%
We propose a mix between trust graphs and economic costs.
%
Rule~\code{4.a} imposes that authors require at least \onerep to post,
effectively blocking Sybil actions.
To vouch for new users, rule~\code{3.a} allows an existing user to like a
newbie's post to unblock it, but at the cost of \onerep.
This cost prevents that malicious members unblock new users indiscriminately,
which would be a breach for Sybils.
For the same reason, rule~\code{2} imposes a temporary cost of \onerep for
each new post.
%
Note that the pioneers rule~\code{1.a} solves the chicken-and-egg problem
imposed by rule~\code{4.a}: if new authors start with no \reps, but require
\reps to operate, it is necessary that some authors have initial \reps to boot
the chains.

@@

{\scriptsize
\begin{verbatim}
 > freechains keys pubpvt 'newbie-password'
 503AB5.. 41DDF1..  <-- public and private keys
 > freechains 'forum' join '4B56AD..'
 10AE3E..           <-- same pioneer as before
 > freechains 'forum' post 'Im a newbie...' \
    --sign='41DDF1..'
 2_C3A40F..         <-- blocked post
 > freechains 'forum' like '2_C3A40F..' \
    --sign='DA3B5F..'
 3_59F3E1..         <-- hash representing the like
\end{verbatim}
}

Note that chains with the same name but different pioneers would be
incompatible because the hash of genesis posts depend on the pioneers' public
keys.

@@

Back to Figure~\ref{fig.forum}, a new user with \nreps{0} tries to post a
message (hash~\code{C3A40F..}) and is blocked (rule~\code{4.a}), as the red
background highlights.
But the pioneer liked the blocked message, decreasing herself to \nreps{29}
and increasing new user to \onerep (rule~\code{3.a}).
Note that the newbie post is not penalized (rule~\code{2}) because it is
followed by the pioneer like, which still controls all \reps in the chain.

With no additional rules to generate \reps, the initial \nreps{30} would
constitute the whole ``chain economy'' forever.
For this reason, rule~\code{1.b} rewards authors of new posts with \onerep,
but only after 24 hours.
This rule stimulates content creation and grows the economy of chains.
The 24-hour period gives sufficient time for other users to judge the post
before rewarding the author.
It also regulates the growth speed of the chain.
In Figure~\ref{fig.forum}, after 1 day, the pioneer would now accumulate
\nreps{30} and the new user \nreps{2}, growing the economy in \nreps{2} as
result of the two consolidated posts.
Note that rule~\code{1.b} rewards at most one post of each author at a time.
Hence, new posts during the 24-hour period will not reward each of them with
extra \reps.
Note also that rule~\code{4.b} limits authors to at most \nreps{30}, which
provides incentives to spend likes and thus decentralize the network, while
rule~\code{4.c} limits the size of posts to at most \emph{128kB} to prevent
DDoS attacks using gigantic blocked posts.

Likes and dislikes (rules \code{3.a} and \code{3.b}) serve three purposes
in the chains:
    (i) welcoming new users,
    (ii) measuring the quality of posts, and
    (iii) revoking abusive posts (SPAM, fake news, illicit content, etc).
%
The quality of posts is subjective and is up to users to judge them with likes,
dislikes, or simply abstaining.
%
This way, access to chains is permissionless in the sense that the actual peers
and identities behind posts are not directly assessed by the protocol, but
instead by the other users in the system.
%
The reputation of a given post is the difference between its likes and
dislikes, which can be used in end-user software for filtering and highlighting
purposes.
%
On the one hand, since \reps are finite, users need to ponder to avoid
indiscriminate expenditure.
On the other hand, since \reps are limited to at most \nreps{30} per author
(rule~\code{4.b}), users also have incentives to rate content.
Hence, these upper and lower limits work together towards the quality of the
chains.
%
Note that a dislike shrinks the chain economy since it removes \reps from both
the origin and target.
Finally, the actual contents of a post may be revoked if it has at least 3
dislikes, and more dislikes than likes (rule~\code{3}).
However, considering that \reps are scarce, dislikes are encouraged to combat
abusive behavior, but not to eliminate divergences of opinion.

\section{Experiments with Public Forums}
\label{sec.evaluation}

In this section, we perform experiments to evaluate the performance and
practicability of the protocol.
%
As detailed further, we measure the following evaluation parameters:
    (a) metadata overhead,
    (b) consensus runtime,
    (c) graph forks, and
    (d) blocked messages.

@@

The simulation spawns $N$ peers, each joining the same chain with the same
arguments.
For each message in the original forum, we
    (i)   set the timestamp of all peers to match the date,
    (ii)  create a pair of keys if the author is new,
    (iii) post the message from a random peer in $N$,
    (iv)  like the post if the author has no reputation, and
    (v)   synchronize the chain with $M$ random peers.
%
Since all messages are part of the original archive, we always perform step
(iv) to unblock messages from newbies, which we use to measure the evaluation
parameter (d).
Step (v) will inevitably create forks in the chain for any \texttt{M<N}, which
we measure in evaluation parameter (c).

@@

Evaluation item (b) accounts the consensus algorithm applied locally at each
chain.
We measured the time to sort posts in a local DAG both for the first time
(without any caches), and incrementally (from stable consensus caches).
%
For the chat chain, it takes \texttt{125s} and \texttt{50ms} for the initial
and incremental sorts respectively, while for the newsgroup chain, it takes
\texttt{100s} and \texttt{70ms}.
%
%Currently, we synchronize the DAGs takes hours because we the synchronization time
%
The incremental sort is limited to 7 days or 100 posts, regardless of the size
of the chain, which conveniently settles an upper bound on the input size of
the consensus algorithm.
%
Given that we use plain JSON files in the file system, we consider an
incremental consensus under \texttt{100ms} to be a practical upper bound.

Evaluation item (c) counts the number of forks in chain DAGs, which indicates
the level of asynchrony between peers following the local-first principle
(supposedly).
We calculate the ratio of forks over the total number of messages, e.g., a DAG
with $100$ messages and $10$ forks has a ratio of $10\%$.
%Based on the choices for $M$ and $N$ for each forum, we expect the newsgroup to
%have a higher ratio, since peers synchronize less.
We found a ratio of $18\%$ for the chat and $14\%$ for the newsgroup, which
confirms that the simulation achieves a reasonable level of asynchrony.

>>>>

\begin{table}
\centering
\includegraphics[width=0.75\textwidth]{general.png}
\caption{General reputation rules in public forums.}
\label{fig.general}
\end{table}

We propose a reputation system in which users can spend tokens named \reps to
post and rate content in the forums:
a \code{post} initially penalizes authors until it consolidates and counts
positively;
a \code{like} is a positive feedback that helps subscribers to distinguish
content amid excess;
a \code{dislike} is a negative feedback that revokes content when crossing a
threshold.
Table~\ref{fig.general} summarizes the general reputation rules and their
goals.
To prevent Sybils, users with no \reps cannot operate under these rules,
requiring a welcoming like from any other user already in the system.
The fact that likes are zero-sum operations, which only transfer reputation,
eliminates the incentives from malicious users to invite Sybils into the
system.
The only way to generate new \reps is to post content that other users
tolerate, which demands non-trivial work resistant to automation.%
\footnote{
    Note that to some extent AI content may be indistinguishable from human
    content, but its frequency and relevance in forums is still subject to
    human evaluation in our proposal.
}

@@

\subsection{Basic Reputation Consensus}
\label{sec.design.basic}

@@

Our key idea to stablish consensus in forum DAGs is to favor forks with posts
from users that constitute the majority of the reputation in the network.
These forks have more associated work from active users and are analogous to
longest chains in Bitcoin.
%
In technical terms, we adapt a topological sorting algorithm to favor
reputation when deciding between branches in a forum DAG.

Figure~\ref{fig.reps}.A illustrates the evolution of a forum DAG in the
presence of a network partition:
A common prefix has signed posts from users $a$, $b$, and $c$, when they were
still all connected.
%
We assume that within the common prefix, users $a$ and $b$ have contributed
with better content and therefore have more reputation combined than $c$ has
alone (i.e., $8+5>3$).
We will use this information when ordering branches in the consensus algorithm.
%
Then, user $c$ disconnects from $a$ and $b$, and evolves in separate along with
new users $x$ and $y$ with posts in \code{branch-1}.
In the meantime, $a$ and $b$ participate with posts in \code{branch-2}.
After some time, the partitions reconnect and new posts from $x$, $b$ and $c$
merge the branches with a common suffix.

Note that even non-malicious network partitions may experience conflicts when
merging branches.
For instance, dislike operations in one branch to common past posts could
possibly block posts from users in the other branch, depending on how they are
ordered.
Therefore, we need to order the branches with a global consensus, such that we
can validate operations consistently across the network.
%
Note also that, to reach consensus, we should preferably rely on the maximum
information that all branches agree on, which is exactly the posts in the
common prefix.

Therefore, based on the reputation of users in the common prefix,
Figure~\ref{fig.reps}.B indicates consensus order between posts in the forum
DAG:
    the common prefix comes first,
    followed by all posts in \code{branch-2},
    followed by all posts in \code{branch-1},
    followed by the common suffix.
%
The \code{branch-2} with users $a$ and $b$ takes priority over \code{branch-1}
with user $c$ because, before the forking point, $a$ and $b$ have more
reputation than $c$, $x$, and $y$ combined.

The basic consensus rule is therefore straightforward:
Whenever a fork is found, we create sets of users with posts in each branch
($B1=\{c,x,y\}$ and $B2=\{a,b\}$).
Then, we take each set, and sum its users \reps based on the common prefix
($S1=3$ and $S2=13$), since it is the maximum information that all branches
agree.
Finally, we order the branches based on the highest sum of \reps found
(Figure~\ref{fig.reps}.B).

While applying the branches in order, if any post operation fails, all
remaining posts are rejected and removed from the DAG, as if they never
existed.
As an example, suppose that the last post by $a$ (in gray) is a dislike to user
$c$.
Then, it's possible that the last post by $c$ (in red), now suppose with
\nreps{0}, is rejected together with all posts in sequence, including those in
the common suffix.
%
This rejection rule for remaining posts is a direct consequence of tamper proof
Merkle~DAGs~\cite{p2p.ipfs,p2p.bitcoin} typically used in permissionless
protocols, which cannot support graph modifications.

There are some other relevant considerations about forks and merges:
%
Peers that first received branches with less reputation need to reorder all
posts starting at the forking point.
This might involve removing content in the end-user software.
This behavior is similar to Bitcoin's blockchain reorganization, when a peer
detects a new longest chain. % and disconsiders old blocks.
%
Likewise, peers that first saw branches with more reputation just need to put
the other branch in sequence with no reordering at all.
%and do not need to recompute anything.
This behavior is expected to happen in the majority of the connected network.
%
Note that unlike Bitcoin, forks are not only permitted but encouraged due to
the local-first software principle~\cite{p2p.local}, in which networked
applications can work locally while offline.
However, the longer a peer remains disconnected, the more conflicting
operations it may see, and the higher are the chances of reordering when
rejoining.

In summary, it is important to remark that
    (i) branches always introduce conflicts, which
    (ii) require consensus to order operations, which
    (iii) is based on reputation of an agreed common prefix,
    (iv) leading to a global and deterministic result.
%
Note that the consensus order exists for reputation accountability purposes,
and is only a view of the primary DAG structure.

\subsection{Malicious Behaviors and Hard Forks}
\label{sec.design.hard}

Now, consider a malicious scenario in which user $c$ in Figure~\ref{fig.reps}
intentionally disconnects from the network for weeks to cultivate fake
identities $x$ and $y$ and accumulate \reps.
When merging, these fake users become legitimate and now accumulate the
majority of \reps in the forum.
%
However, this kind of attack is refutable because, since the common prefix is
the basis of consensus, users in the branch with more reputation can still
react even a posteriori.
For instance, users $a$ and $b$ can pretend that they did not yet see
\code{branch-1} and post extra dislikes to user $c$ in \code{branch-2} so that
a further merge invalidates all posts from \code{branch-1}, removing them from
the DAG.
%
This kind of attack is innocuous because it requires at least half of the \reps
accumulated in the common prefix.
%
The same countermeasure can be used against a minority of malicious users
colluding to boost or protect their own posts.

@@

As a counterpoint, maybe users $a$ and $b$ have abandoned the forum for months,
and thus \code{branch-1} is actually legitimate.
In this case, users $a$ and $b$ might be the ones trying to take over the
forum and would succeed, since they control the majority of \reps in the common
prefix.
Yet another possibility is that both branches are legitimate but became
disconnected for a long period of time.
In any case, it is unacceptable for correct users that a very old remote branch
affects a long active forum.

For this reason, the consensus algorithm includes an extra constraint that
prevents long-lasting local branches to merge, which leads to \emph{hard forks}
in the network.
A hard fork occurs when a local branch crosses a predetermined and irreversible
threshold of \emph{7 days} or \emph{100 posts} of activity.
In this case, regardless of the remote branch reputation, the local branch
takes priority and is ordered first.
This situation is analogous to a hard fork in Bitcoin and the branches become
incompatible and will never synchronize again.
More than simple numeric disputes, a hard fork represents a social conflict in
which reconciling branches is no longer possible.

Figure~\ref{fig.hard} illustrates hard forks by distinguishing \emph{stable
consensus}, which cannot be reordered, from \emph{unstable consensus}, which
may still be affected by incoming branches.
The activity threshold counts backwards, starting from the newest local post
in the unstable consensus to older posts until they are permanently frozen.

In summary, the rules to merge a branch from a remote machine \code{j} into a
branch from a local machine \code{i} are as follows:
\begin{itemize}
    \item \code{i} is first if it crosses the activity threshold of
          \emph{7 days} or \emph{100 posts}, regardless of \code{j}.
    \item \code{i} or \code{j} is first, whichever has more reputation in the
          common prefix, as discussed in Section~\ref{sec.design.basic}.
    \item otherwise, branches are ordered by an arbitrary criteria, such as
          lexicographical order of the post hashes immediately after the
          common prefix.
\end{itemize}

To conclude this section, we should keep in mind that our consensus algorithm
is based on subjective evaluations of users and posts.
For this reason, unlike Bitcoin, in which forks are discouraged and a single
blockchain must prevail, public forums can split and diverge indiscriminately.
For instance, users can explicitly apply hard forks at any point in time to
``reboot'' a community from a previous state.
Ultimately, the consensus algorithm provides a transparent mechanism to help
users understand the evolution of public forums and act accordingly, even if it
leads to hard forks.

@@

Finally, we consider that revoking posts is fundamental in the context of
social content publishing, and thus, an important contribution of this work.

As described in Figure~\ref{fig.state}, a forum post has three possible states:
\code{BLOCKED}, \code{ACCEPTED}, or \code{REVOKED}.
%
If the author has reputation, a new post is immediately \code{ACCEPTED} in the
forum.
%
Otherwise, it is \code{BLOCKED} and requires a like from another user.

Blocked posts are not considered part of the forum DAG in the sense that new
posts do not link back to it.
In addition, peers are not required to hold blocked posts and neither
retransmit them to other peers.
However, if blocked posts are not disseminated, new users will never have the
chance to be welcomed with a like.
A reasonable policy is to hold blocked posts in a temporary but public bag and
retransmit them for some visibility in the network.

Once accepted, a post becomes part of the forum and can never be removed
again, since Merkle~DAGs are immutable by design.
%Note that blocked posts that become accepted are always succeeded by a
%\emph{like} (Figure~\ref{fig.forum}).
%
However, if the number of dislikes exceeds an arbitrary threshold (e.g., more
dislikes than likes), the post becomes \code{REVOKED} and its payload is not
retransmitted to other peers.
Note that a post hash does not depend on its associated payload, but only on
the payload hash.
Hence, it is safe to remove the payload as long as one can prove its revoked
state.
Later, if the post receives new likes, it means that the payload is still known
somewhere and peers can request it when synchronizing again, since the post
becomes \code{ACCEPTED} again.

\subsection{Peer Synchronization and Byzantine Faults}

Similarly to Bitcoin, the underlying network protocol replicates on each peer
the full state of the Merkle-DAG representing the causal relationships between
the messages.
When peers synchronize, they mutually exchange missing branches such that they
reach the same DAG structure~\cite{p2p.byz}.
As they synchronize, each peer runs the consensus algorithm locally and
verifies if all received posts are consistent.

Because the protocol relies on a tamper-proof DAGs, and because each peer
self verifies the posts as they synchronize, the protocol becomes tolerant to
Byzantine faults~\cite{lamport.byz}.
%
Note that no information coming from any peer is trusted a priori.
%
For instance, a malicious peer that tries to synchronize branches with erratic
data can be detected on the first inconsistent post.
This allows the correct peer to close the connection immediately and blacklist
the malicious peer from future connections.

Therefore, although Byzantine peers may still perform some sort of denial of
service attacks in the network, they cannot modify the DAG structure, create
arbitrary content (without performing work), or impede that two correct nodes
synchronize directly.

@@

We integrated the proposed reputation and consensus mechanism with \FC to
support content moderation and mitigate abuse.
Table~\ref{fig.rules} details the concrete reputation rules we conceived for
\FC, which is compatible with the general rules of Table~\ref{fig.general}, and
are discussed as follows through an example.
Authors have to sign posts in order to be accounted by the reputation system
and operate in the chains.
We start by creating an identity whose public key is assigned as the pioneer in
a chain named \emph{forum}:

{\footnotesize
\begin{verbatim}
 > freechains keys pubpvt 'pioneer-password'
 4B56AD.. DA3B5F..  <-- public and private keys
 > freechains 'forum' join '4B56AD..'
 10AE3E..           <-- hash representing the chain
 > freechains 'forum' post --sign='DA3B5F..' 'The chain purpose is...'
 1_CC2184..         <-- hash representing the post
\end{verbatim}
}

The \code{join} command in rule~\code{1.a} bootstraps a public chain,
assigning \nreps{30} equally distributed between an arbitrary number of
pioneers indicated through their public keys (one in this example).
The pioneers shape the initial culture of the chain with the first posts and
likes, while they gradually transfer \reps to other authors, which also
transfer to other authors, expanding the community.
%
The \code{post} command in sequence is signed by the single pioneer and
indicates the purpose of the chain for future users.

The most basic concern in public forums is to resist Sybils abusing the
chains.
Fully peer-to-peer systems cannot rely on logins or {\footnotesize CAPTCHAs}
due to the lack of a central authority.
Viable alternatives include (i) building social trust graphs, in which users
already in the community vouch for new users, or (ii) imposing explicit costs
for new posts, such as proof of work.
%
We propose a mix between trust graphs and economic costs.
%
Rule~\code{4.a} imposes that authors require at least \onerep to post,
effectively blocking Sybil actions.
To vouch for new users, rule~\code{3.a} allows an existing user to like a
newbie's post to unblock it, but at the cost of \onerep.
This cost prevents that malicious members unblock new users indiscriminately,
which would be a breach for Sybils.
For the same reason, rule~\code{2} imposes a temporary cost of \onerep for
each new post.
%
Note that the pioneers rule~\code{1.a} solves the chicken-and-egg problem
imposed by rule~\code{4.a}: if new authors start with no \reps, but require
\reps to operate, it is necessary that some authors have initial \reps to boot
the chains.

@@

{\footnotesize
\begin{verbatim}
 > freechains keys pubpvt 'newbie-password'
 503AB5.. 41DDF1..  <-- public and private keys
 > freechains 'forum' join '4B56AD..'
 10AE3E..           <-- same pioneer as before
 > freechains 'forum' post 'Im a newbie...' --sign='41DDF1..'
 2_C3A40F..         <-- blocked post
 > freechains 'forum' like '2_C3A40F..' --sign='DA3B5F..'
 3_59F3E1..         <-- hash representing the like
\end{verbatim}
}

Note that chains with the same name but different pioneers would be
incompatible because the hash of genesis posts depends on the pioneers' public
keys.

@@

Back to Figure~\ref{fig.forum}, a new user with \nreps{0} tries to post a
message (hash~\code{2\_C3A40F..}) and is blocked (rule~\code{4.a}), as the red
background highlights.
But the pioneer likes the blocked message, decreasing herself to \nreps{29}
and increasing new user to \onerep (rule~\code{3.a}).
Note that the newbie post is not penalized (rule~\code{2}) because it is
followed by the pioneer like, which still controls all \reps in the chain.

With no additional rules to generate \reps, the initial \nreps{30} would
constitute the whole ``chain economy'' forever.
For this reason, rule~\code{1.b} rewards authors of new posts with \onerep,
but only after 24 hours.
This rule stimulates content creation and grows the economy of chains.
The 24-hour period gives sufficient time for other users to judge the post
before rewarding the author.
It also regulates the growth speed of the chain.
In Figure~\ref{fig.forum}, after 1 day, the pioneer would now accumulate
\nreps{30} and the new user \nreps{2}, growing the economy in \nreps{2} as
result of the two new consolidated posts.
Note that rule~\code{1.b} rewards at most one post of each author at a time.
Hence, extra posts during the 24-hour period do not reward authors with extra
\reps.
Note also that rule~\code{4.b} limits authors to at most \nreps{30}, which
provides incentives to spend likes and thus decentralize the network, while
rule~\code{4.c} limits the size of posts to at most \emph{128kB} to prevent
DDoS attacks using gigantic blocked posts.

Likes and dislikes (rules \code{3.a} and \code{3.b}) serve three purposes
in the chains:
    (i) welcoming new users,
    (ii) measuring the quality of posts, and
    (iii) revoking abusive posts (SPAM, fake news, illicit content, etc).
%
The quality of posts is subjective and is up to users to judge them with likes,
dislikes, or simply abstaining.
%
This way, access to chains is permissionless in the sense that the actual peers
and identities behind posts are not directly assessed by the protocol, but
instead by the judgement of other users in the system.
%
The reputation of a given post is the difference between its likes and
dislikes, which can be used in end-user software for filtering and highlighting
purposes.
%
On the one hand, since \reps are finite, users need to ponder to avoid
indiscriminate expenditure.
On the other hand, since \reps are limited to at most \nreps{30} per author
(rule~\code{4.b}), users also have incentives to rate content.
Hence, these upper and lower limits work together towards the quality of the
chains.
%
Note that a dislike shrinks the chain economy since it removes \reps from both
the origin and target.
Finally, the actual contents of a post may be revoked if it has at least 3
dislikes, and more dislikes than likes (rule~\code{3}).
However, considering that \reps are scarce, dislikes are encouraged to combat
abusive behavior, but not to eliminate divergences of opinion.

\subsection{Experiments with Public Forums in \FC}
\label{sec.evaluation}

In this section, we perform experiments to evaluate the performance and
practicability of \FC and its consensus algorithm.
%
As detailed further, we measure the following evaluation parameters:
    (a) metadata overhead,
    (b) consensus runtime,
    (c) graph forks, and
    (d) blocked messages.

@@

The simulation spawns $N$ peers, each joining the same chain with the same
arguments.
For each message in the original forum, we
    (i)   set the timestamp of all peers to match the original date,
    (ii)  create a pair of keys if the author is new,
    (iii) post the message from a random peer in $N$,
    (iv)  like the post if the author has no reputation, and
    (v)   synchronize the chain with $M$ random peers.
%
Since all messages are part of the original archive, we always perform step
(iv) to unblock messages from newbies, which we also use to measure the
evaluation parameter (d).
Step (v) will inevitably create forks in the chain for any \texttt{M<N}, which
we measure in evaluation parameter (c).

@@

Evaluation item (b) accounts the consensus algorithm applied locally at each
chain.
We measured the time to sort posts in a local DAG both for the first time
(without any caches), and incrementally (from stable consensus caches).
%
For the chat chain, it takes \texttt{125s} and \texttt{50ms} for the initial
and incremental sorts respectively, while for the newsgroup chain, it takes
\texttt{100s} and \texttt{70ms}.
%
%Currently, we synchronize the DAGs takes hours because we the synchronization time
%
The incremental sort is limited to 7 days or 100 posts by design (as discussed
in Section~\ref{sec.design.hard}), regardless of the size of the chain, which
conveniently settles an upper bound on the input size of the consensus
algorithm.
%
Given that we use plain JSON files in the file system, we consider an
incremental consensus under \texttt{100ms} to be a practical upper bound.

Evaluation item (c) counts the number of forks in chain DAGs, which indicates
the level of asynchrony between peers following the local-first principle.
We calculate the ratio of forks over the total number of messages, e.g., a DAG
with $100$ messages and $10$ forks has a ratio of $10\%$.
%Based on the choices for $M$ and $N$ for each forum, we expect the newsgroup to
%have a higher ratio, since peers synchronize less.
We found a ratio of $18\%$ for the chat and $14\%$ for the newsgroup, which
confirms that the simulation achieves a reasonable level of asynchrony.
